---
title: "2022.10.18_Rclass_Practice5.1"
author: "Berrica"
date: "2022-10-24"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Interactive Map
Originally, I want to make a map containing both global coral distribution and diving sites to see the relationship between them.
However, it is difficult to find free GPS or GIS data of global coral distribution and diving sites.
Therefore, I found the data from Florida instead of global data.
I marked the location of my house to see how these diving site is far from my house.
```{r}
#Load packages
library(leaflet)
library(rgbif)
library(tidyverse)
library(readxl)

#Read data from Florida
data <- read_xlsx('Data/diving_site.xlsx')

#data sources: https://www.scubadiversworld.com/sites.html

#Make icon to mark my house: diver icon

diver.icon <- makeIcon(
  iconUrl = "https://cdn2.iconfinder.com/data/icons/sport-ii/100/24-512.png",
  iconWidth = 18, iconHeight = 20)


#Make different icon for wreck diving and reef diving
icons <- iconList(Wreck = makeIcon('https://www.freeiconspng.com/thumbs/shipping-icon/-nautical-raw-ship-shipping-simple-yacht-icon--icon-search-engine-3.png', iconWidth = 18, iconHeight =15),
                  Reef = makeIcon('https://cdn-icons-png.flaticon.com/512/202/202206.png', iconWidth = 18, iconHeight =18)) #iconList can be used to designed different size of icon

#Draw Map
M <- leaflet() %>% 
  addTiles() %>% setView(30.664237, 24.258327, zoom = 1) %>%   #set the beginning map center 
  addMarkers(lng=120.664237, lat = 24.258327, popup= "My sweet home", icon =  diver.icon) %>%  #mark my house with diver icon
  addMarkers(~Lon, ~Lat, data = data, icon = icons)  #mark different types of diving with corresponding icons
M

```

## Welcome to Florida! Enjoy the underwater world here!

If you want to go reef diving, check the location of those coral icons on the map!
If you want to go wreck diving, check the location of those icons on the map!

```{r pressure, echo=FALSE}
M
```


