---
title: "2022.10.04_R_class_Practice3.2"
author: "Berrica"
date: "2022-10-04"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Practice3.2 in R class

```{r, eval = T}
#write the original dataframe
subject <- paste('subject', '_', 1:5)
bubble.tea <- data.frame(before_diet = c(104, 95, 87, 77, 112), 
                         after_diet = c(96, 91, 81, 75, 118),
                         row.names = subject)
mode(bubble.tea)
str(bubble.tea)
bubble.tea
```

## The first element
```{r, eval = T}
#The first element----
#call pakages
library(tidyr)
library(dplyr)

#reformat the last dataframe to obtain column 'weight' and column 'time'
first <- bubble.tea %>% pivot_longer(cols = before_diet:after_diet, 
                                     names_to = 'time', 
                                     values_to = 'weight')
mode(first)
str(first)
first
```

## The second element
```{r, eval = T}

#[1]the row names of the table previously created
rn <- row.names(bubble.tea) #it is a vector of character
mode(rn)
rn

#[2]indicate the weight loss (in %) of each subject 
weight_loss <- (bubble.tea$before_diet - bubble.tea$after_diet)/bubble.tea$before_diet*100 #it's a numeric vector (double)
mode(weight_loss)
typeof(weight_loss)
weight_loss

#[3]combine these two elements in a table with two columns: subject and weight_loss
combine <- data.frame(subject = rn,
                      weight_loss = weight_loss)
mode(combine)
str(combine)
combine

WEIGHT_LOSS <- list(rn, weight_loss, combine)
WEIGHT_LOSS
```

## The third element: How much I enjoy manipulating data in R ?
```{r, eval = T, echo = F}
# install.packages("ggplot2")
# install.packages("pryr")

library(ggplot2)
library(pryr)

# heart curve formula
heart <- quote((x^2 + y^2 - 1)^3 - x^2 * y^3)

# formula for heart curve at a given x
heart_at_x <- function(x) {
  function(y) eval(substitute_q(heart, list(x = x)), list(y = y))
}

# trace the heart curve
# by evaluating the heart curve formula at each x, then finding the roots of the
# resulting formula in y; e.g. a x==0, find the roots of (y^2 - 1)^3 = 0
# broken up into upper and lower parts (h_y1 and h_y2)
heart_x <- seq(-1.136, 1.136, 0.001)
heart_y_lower <- sapply(heart_x, function(x) uniroot(heart_at_x(x), c(-2, 0.6))$root)
heart_y_upper <- sapply(heart_x, function(x) uniroot(heart_at_x(x), c(0.6, 2))$root)

# put together data frame
heart_df <- data.frame(x = rep(heart_x, 2), 
                       y = c(heart_y_lower, heart_y_upper))

# show outline
#with(heart_df, plot(x, y))

# create a data frame with one row per x, so we can fill in the heart
heart_df_minmax <- data.frame(x = heart_x,  
                              y_min = heart_y_lower, 
                              y_max = heart_y_upper)

set.seed(20150214)

# fill in the heart by generating random deviates at each x 
# and rejecting those that fall outside the heart curve
heart_full <- apply(heart_df_minmax, 
                    1, 
                    function(w) {
                      x <- w["x"]
                      y_min = w["y_min"]
                      y_max = w["y_max"]
                      y <- rnorm(2, mean = 0.33)
                      y <- y[between(y, y_min, y_max)]
                      x <- x[any(is.finite(y))]
                      data.frame(x, y, row.names = NULL)
                    })

# change from list to data frame
heart_full <- bind_rows(heart_full)

# add random numbers for color and size
heart_full <- heart_full %>% 
  mutate(z1 = runif(n()), 
         z2 = pmin(abs(rnorm(n())), 3), 
         order = runif(n())) %>%
  arrange(order)


# plot the heart
p <- ggplot(heart_full, 
            aes(x, y, color = z1, size = z2)) + 
  geom_point(pch = -1 * as.hexmode(9829)) + 
  scale_color_gradient(limits = c(0, 1), low = "red3", high = "pink") + 
  scale_size(limits = c(0, 3), range = c(0.1, 20)) + 
  theme_bw() +
  theme(legend.position = "none", 
        axis.title.x = element_blank(),
        axis.text.x = element_blank(), 
        axis.title.y = element_blank(),
        axis.text.y = element_blank(), 
        axis.ticks = element_blank(),
        panel.border = element_blank(),
        plot.title = element_text(face = 'bold', size = 60, hjust = 0.5, vjust = -12))+
  scale_y_continuous(breaks = NULL) +
  scale_x_continuous(breaks = NULL) +
  ggtitle("I Love R")
p

```

### Manipulating data in R is quite interesting. When I am manipulating R, it is like **playing** R. Endless surprise will make you be crazy for it!
### In R, there are a lot of different functions that can help me calculate and visualize data. In brief, it is really useful. Welcome to enjoy R with us! 


## Outcome
```{r, eval = T}
#Make a list called BUBBLE_DIET
BUBBLE_DIET <- list(first, WEIGHT_LOSS, ILOVER = p)
mode(BUBBLE_DIET)
str(BUBBLE_DIET)
```
